---
title: "Regression Models Course Project"
author: "MsGret"
date: "September 25, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Based on the data set of a collection of cars (`mtcars` data set), we explore the relationship between a set of variables and miles per gallon (MPG) (outcome) and answer two questions:

- “Is an automatic or manual transmission better for MPG”
- "Quantify the MPG difference between automatic and manual transmissions"

## Exploratory Analyses

Load the data and perform some basic exploratory data analyses

```{r data}
library (datasets)
data(mtcars)
dim(mtcars)
str(mtcars)
```

Data set consists consists of 11 variables and 32 observation for each variable.

Look at relationship between transmission type (`am` as factor variable (0 - automatic, 1 - manual) and miles per gallon (`mpg`) (**Appendix A**).

```{r}
mtcars$am <- factor(mtcars$am, labels = c("automatic", "manual"))
```

Based on boxplot in **Appendix A** we can suppose that there is a significant 
difference in MPG for different transmission type.


## Statistical Inference

Test our hypothesis: Null hypothesis is "the MPG means for different transmission type is equal" or "true difference in MPG means for different transmission type is equal to 0".

```{r}
t.test(mpg ~ am, data = mtcars)
```

We can reject null hypothesis that the difference in MPG means for different transmission type is equal to 0 (p-value = 0.001374).

## Regression Analysis

So MPG depends on transmission type, but define how other variables affect on MPG. Build multivariable regression model (results in **Appendix B**):

```{r fullModel,results='hide'}
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs, labels = c("V", "S"))
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)

fullModel <- lm(mpg ~ ., data = mtcars)
summary(fullModel)
```

So none of the coefficients have a p-value less than 0.05 (statistically significant). Find better model (based on removing variables from the model and evaluating the AIC):

```{r,results='hide'}
AICModel <- step(fullModel, direction = "both")
summary(AICModel)
anova(AICModel, fullModel)
```

Comparing the `AICModel` with the `fullModel` we see that removing other predictors has not significantly affected the explanatory ability of the model.

The `AICModel` explains about 87% of the variance in MPG (R-squared is 0.8659). The coefficients conclude that increasing the number of cylinders from 4 to 6 with decrease the MPG by 3.031, but from 4 to 8 with decrease the MPG by 2.164. One additional horsepower is decreases MPG by 0.0321. Weight decreases the MPG by 2.497 for each 1000 lbs increase. A Manual transmission improves the MPG by
1.809.

## Residual Analysis

Based on residuals plots (**Appendix C**) we can conclude:

- the `Residuals vs Fitted` plot doesn't show pattern and confirms that residuals are independent;
- the `Normal Q-Q` plot confirms that the residuals are normally distributed (with some deviate from normality at the tails);
- the `Scale-Location` confirms the constant variance assumption;
- the `Residuals vs Leverage` confirms that there are no outliers (all values fall within the 0.5 bands).

## Conclusion

- There is a significant difference in MPG for different transmission type (MPG mean for manual type more automatic type at 7.24).
- Based on `AICModel` we can conclude that number of cylinders, weight and horsepower are more statistically significant then transmission type for determining MPG.


## Appendix A
 
```{r echo=FALSE}
  boxplot(mpg ~ am, data = mtcars,
          xlab="Transmission Type", 
          ylab="Miles per Gallon",
          main="MPG vs. Transmission Type")
```

## Appendix B

```{r}
summary(fullModel)
summary(AICModel)
anova(AICModel, fullModel)
```

## Appendix C

Residuals plots for `AICModel`:

```{r}
par(mfrow = c(2, 2))
plot(AICModel)
```
